M.Sc. Proposal: \\ \bigskip Computational Electrokinetics

### 2011-11-29 notes/summary/ch01.tex
Discretization notes.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Microscale}
Denote positive and negative ionic concentration by $c_+$ and $c_-$ respectively.
The respective fluxes are denoted by $\bj_+$ and $\bj_-$.

@@

Due to cation selectivity, $c_+|_{\rho=0} = \gamma$.
Denote zeta potential as the voltage drop between the particle surface and the end of
the boundary layer, $\zeta = \cV - \varPhi$. 
Without loss of generality, 
we assume that $\cV = -\log\gamma$, thus $\zeta = -\log\gamma-\varPhi$:
\begin{eqnarray}
\varPhi &=& -\log C
\end{eqnarray}

@@

Leading-order tangential momentum balance in the Debye layer yields:
\begin{eqnarray} \nonumber
\deriv{^2 v_\theta}{\rho^2} - \deriv{p}{\theta} 
 + \deriv{\varphi}{\theta} \deriv{^2\varphi}{\rho^2} &=& 0
\end{eqnarray}
By integrating from $\rho = 0$ to $r \rightarrow \infty$, we have Dukhin-Derjaguin slip formula
for tangential velocity component $V_\theta$ (where as the radial component is $V_r = 0$):
\begin{eqnarray}
V_\theta &=& \zeta \cdot \deriv{\varPhi}{\theta} + 
      2\log\pars{1 - \tanh^2 \frac{\zeta}{4}} \cdot \deriv{}{\theta} \log C 
 = \zeta \cdot \deriv{\varPhi}{\theta} -
      4\log\cosh \frac{\zeta}{4} \cdot \deriv{}{\theta} \log C
\end{eqnarray}
where $\zeta = \cV - \varPhi = -\log\gamma - \varPhi$.

Note that due to boundary condition, $\log C = -\varPhi$, thus:
\begin{eqnarray}
V_\theta &=& \zeta \cdot \deriv{\varPhi}{\theta} + 
      4\log\cosh \frac{\zeta}{4} \cdot \deriv{\varPhi}{\theta} = 
4\log\pars{\frac{1 + \exp\left\{\zeta/2\right\}}{2}} \cdot \deriv{\zeta}{\theta}
\end{eqnarray}

\subsection{Steady state}
The stress tensor is composed of Newtonian and Maxwel stresses:
\begin{eqnarray}
\tT &=& \bnabla \bV + (\bnabla \bV)^\dagger - P \tI
+ \bnabla \varPhi \bnabla \varPhi - \frac{1}{2} (\bnabla \varPhi \cdot \bnabla \varPhi) \tI \\
\bnabla \cdot \tT &=& \bzero
\end{eqnarray}
The total force acting on the particle is computed by the following surface integral:
\begin{eqnarray}
\bF &=& \oint_\mathcal{S} \tT \cdot \bnhat  dA 
\end{eqnarray}
In steady state (constant particle drift velocity), it should be zero: $\bF = \bzero$.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Micro-scale equation}
Denote positive and negative ionic concentration by $c_+$ and $c_-$ respectively.
The respective fluxes are denoted by $\bj_+$ and $\bj_-$.

@@

Due to cation selectivity, $c_+|_{\rho=0} = \gamma$.
\begin{eqnarray}
 \nonumber c_+ = \gamma &=& C \exp(\varPhi - \cV) \\
 \nonumber \cV + \log \gamma &=& \log C + \varPhi
\end{eqnarray}

Denote zeta potential as the voltage drop between the particle surface and the end of
the boundary layer, $\zeta = \cV - \varPhi$. 
Without loss of generality, 
we assume that $\cV = -\log\gamma$, thus $\zeta = -\log\gamma-\varPhi$:
\begin{eqnarray}
\varPhi &=& -\log C
\end{eqnarray}

@@

Leading-order tangential momentum balance in the Debye layer yields:
\begin{eqnarray} \nonumber
\deriv{^2 v_\theta}{\rho^2} - \deriv{p}{\theta} 
 + \deriv{\varphi}{\theta} \deriv{^2\varphi}{\rho^2} &=& 0
\end{eqnarray}
By integrating from $\rho = 0$ to $r \rightarrow \infty$, we have Dukhin-Derjaguin slip formula
for tangential velocity component $V_\theta$ (where as the radial component is $V_r = 0$):
\begin{eqnarray} \nonumber
V_\theta &=& \zeta \cdot \deriv{\varPhi}{\theta} + 
      2\log\pars{1 - \tanh^2 \frac{\zeta}{4}} \cdot \deriv{}{\theta} \log C 
 = \zeta \cdot \deriv{\varPhi}{\theta} -
      4\log\cosh \frac{\zeta}{4} \cdot \deriv{}{\theta} \log C
\end{eqnarray}
where $\zeta = \cV - \varPhi = -\log\gamma - \varPhi$.

Note that due to boundary condition, $\log C = -\varPhi$, thus:
\begin{eqnarray} \nonumber
V_\theta &=& \zeta \cdot \deriv{\varPhi}{\theta} + 
      4\log\cosh \frac{\zeta}{4} \cdot \deriv{\varPhi}{\theta} = 
4\log\pars{\frac{1 + \exp\left\{\zeta/2\right\}}{2}} \cdot \deriv{\zeta}{\theta}
\end{eqnarray}

Since $V_r = 0$ on the boundary, the slip condition can be written as:
\begin{eqnarray}
\bV &=& 
4\log\pars{\frac{1 + \exp\left\{\zeta/2\right\}}{2}} \cdot \bnabla_S \zeta
\end{eqnarray}

\subsection{Steady state}
The stress tensor is composed of Newtonian and Maxwel stresses:
\begin{eqnarray}
\tT &=& \bnabla \bV + (\bnabla \bV)^\dagger - P \tI
+ \bnabla \varPhi \bnabla \varPhi - \frac{1}{2} (\bnabla \varPhi \cdot \bnabla \varPhi) \tI \\
\bnabla \cdot \tT &=& \bzero
\end{eqnarray}
The total force acting on the particle is computed by the following surface integral:
\begin{eqnarray}
\bF &=& \oint_\mathcal{S} \tT \cdot \bnhat  dA 
\end{eqnarray}
In steady state (constant particle drift velocity $\cU$), it should be zero: $\bF = \bzero$.

### 2011-11-29 notes/summary/ch02.tex
Discretization notes.

\section{Numerics}
\subsection{Discretization}
\subsubsection{Spherical Coordinates}
Since the system has axial symmetry, the operators are written in 
spherical coordinates $(r,\theta,\phi)$:
\begin{eqnarray}
x &=& r \sin \theta \cos \phi \\
y &=& r \sin \theta \sin \phi \\
z &=& r \cos \theta
\end{eqnarray}
Due to axial symmetry, $\deriv{a}{\phi} = 0$ and $\bphihat \cdot \bA = 0$, 
so the differential operators have the following form:
\subsubsection{Grids choice}
\subsubsection{Operator Representation}
\subsubsection{Boundary Condition}
\subsection{Solver}
\subsubsection{Newton's Method}
\subsubsection{Steady-state}
\subsubsection{Continuation}

>>>>

\section{Numerics}
\subsection{Discretization}
\subsubsection{Spherical Coordinates}
Since the system has axial symmetry, the operators are written in 
spherical coordinates $(r,\theta,\phi)$:
\begin{eqnarray}
\brc{l}{
x = r \sin \theta \cos \phi \\
y = r \sin \theta \sin \phi \\
z = r \cos \theta} 
\end{eqnarray}
Due to axial symmetry, 
so the scalar gradient and the divergence can be written as:
\begin{eqnarray}
\bnabla f &=& \deriv{f}{r}\brhat + \frac{1}{r}\deriv{f}{\theta}\bthetahat \\
\bnabla \cdot \bF &=& \frac{1}{r^2}\deriv{}{r}\pars{F_r r^2 } + 
                     \frac{1}{r \sin\theta}\deriv{}{\theta}\pars{F_\theta \sin\theta}
\end{eqnarray}
The scalar laplacian can be written as:
\begin{eqnarray}
\Laplacian f &=& \frac{1}{r^2}\deriv{}{r}\pars{r^2 \deriv{f}{r}} + 
                     \frac{1}{r^2 \sin\theta}\deriv{}{\theta}\pars{\sin\theta \deriv{f}{\theta}}
\end{eqnarray}
The vector gradient can be written as:
\begin{eqnarray}
\bnabla \bF &=& \deriv{F_r}{r} \brhat \brhat + \deriv{F_\theta}{r} \brhat \bthetahat + 
\frac{1}{r}\pars{\deriv{F_r}{\theta} - F_\theta} \bthetahat \brhat + 
\frac{1}{r}\pars{\deriv{F_\theta}{\theta} + F_r} \bthetahat \bthetahat
\end{eqnarray}

The vector laplacian can be written as:
\begin{eqnarray}
\bLaplacian \bF &=& 
\left(\Laplacian F_r - \frac{2F_r}{r^2} - 
\frac{2}{r^2 \sin\theta} \deriv{\left(F_\theta \sin\theta \right)}{\theta}\right)\brhat
+ \left(\Laplacian F_\theta - \frac{F_\theta}{r^2 \sin^2\theta} + 
\frac{2}{r^2}\deriv{F_r}{\theta}\right) \bthetahat
\end{eqnarray}

\subsubsection{Grids choice}
In order to discretize the differential operators, 
a regular grid of size $N_r \times N_\theta$ is used:
\begin{eqnarray}
(r_i,\theta_j) &\in& [1, \infty) \times [0,\pi] \\ 
r_i &=& (1+\Delta_r)^i \\
\theta_j &=& \Delta_\theta \cdot j
\end{eqnarray}
where a logarithmic grid is used for $r$ and an uniform one is used for $\theta$.
Note that $r_0 = 1$ and $R_{max} = (1+\Delta_r)^{N_r}$. Thus:
\begin{eqnarray}
\Delta_r &=& \pars{R_{max}} ^ \frac{1}{N_r} - 1 \\
\Delta_\theta &=& \frac{\pi}{N_\theta}
\end{eqnarray}

This grid induces disjoint subdivision of the domain 
$\Omega = [1, R_{max}] \times [0,\pi]$ into cells:
\begin{eqnarray}
\bigcup_{ij}\Omega_{ij} &=& \Omega
\end{eqnarray}
A specific cell $\Omega_{ij}$ and its center $(\bar{r}_i, \bar{\theta}_j)$ are defined by:
\begin{eqnarray}
\Omega_{ij} &=& [r_{i-1}, r_{i}] \times [\theta_{i-1}, \theta_{i}] \\
\bar{r}_i &=& \frac{r_{i-1} + r_{i}}{2} \\
\bar{\theta}_i &=& \frac{\theta_{i-1} + \theta_{i}}{2}
\end{eqnarray}

Each variable is discretized with respect of the cell:
\begin{itemize}
\item $\varPhi$, $C$ and $P$ are represented by their value at the center of each cell, 
using a central grid:
\begin{eqnarray}
\varPhi[i,j] &=& \varPhi(\bar{r}_i, \bar{\theta}_j) \\
C[i,j] &=& C(\bar{r}_i, \bar{\theta}_j) \\
P[i,j] &=& P(\bar{r}_i, \bar{\theta}_j) 
\end{eqnarray}
\item $\bV$ is represented by its values at cell boundaries, using a staggered grid:
\begin{eqnarray}
V_r[i,j] &=& V_r(r_i, \bar{\theta}_j) \\
V_\theta[i,j] &=& V_\theta(\bar{r}_i, {\theta}_j)
\end{eqnarray}
\end{itemize}

\subsubsection{Boundary Conditions}
The boundary conditions far away from the ion-exchange boundary ($r\rightarrow\infty$) are:
\begin{eqnarray}
\deriv{\varPhi}{r} &\rightarrow& -\beta \\
C &\rightarrow& 1 \\
V_r &\rightarrow& -\cU \cos\theta \\
V_\theta &\rightarrow& \cU \sin\theta
\end{eqnarray}
The boundary conditions at the ion-exchange boundary ($r=1$) are:
\begin{eqnarray}
\varPhi &=& -\log C \\
\deriv{\varPhi}{r} &=& \deriv{}{r} \log C \\
V_r &=& 0 \\
V_\theta &=& 4\log\pars{\frac{1 + \exp\left\{\zeta/2\right\}}{2}} \cdot \deriv{\zeta}{\theta} \\
 \nonumber \zeta &=& \cV - \varPhi = \log C - \log \gamma
\end{eqnarray}

\subsection{Solver}
\subsubsection{Newton's Method}
\subsubsection{Steady-state}
\subsubsection{Continuation}
